---
editor_options: 
  markdown: 
    wrap: 72
---

2024-05-22 Tumor explants (PMED) RNAseq Analysis from counts
(protein-coding genes only)

############ SET UP

```{r}
library(edgeR) # You will need to install these packages the first time
library(limma)

# input.dir <- ("/Users/taurakx/Desktop/Practice/Exercise_explant_data.Rmd")
# setwd(input.dir)
```

#### 1) Examine the data

```{r}
rna_counts <- read.csv("rna_counts.csv", row.names=1)
sample_meta <- read.csv("sample_metadata.csv", row.names=1)
gene_features <- read.csv("genefeature_data.csv", row.names=1)
```

Exercise questions: Can you write some code to tell how many samples and
how many genes there are in this data? (Hint: they correspond to number
of columns/ rows of some of the tables)

```{r}
nrow(rna_counts) # num genes
ncol(rna_counts) # num samples
```

Can you use "str" function to make a quick summary of all 3 tables given
above?

```{r}
str(gene_features)
str(rna_counts)
str(sample_meta)
```

Can you check if the rownames of sample_meta and colnames of rna_counts
are exactly the same?

```{r}
all(colnames(rna_counts) == rownames(sample_meta))
```

Can you check if the rownames of rna_counts are exactly the same as
"ensembl_gene_id" column in gene_features?

```{r}
all(rownames(rna_counts) == gene_features$ensembl_gene_id)
```

in sample_meta, can you subset out only the samples where indication =
"CRC" (and save it as another dataframe)?

```{r}
CRC_sample_meta <- sample_meta[sample_meta$indication == "CRC",]
```

in sample_meta, can you subset out only the samples where indication =
"CRC" OR "HNSCC"?

```{r}
sample_meta[sample_meta$indication %in% c("CRC", "HNSCC"),]
```

in sample_meta, can you subset out only the samples where indication =
"CRC" OR "HNSCC" OR "BLCA" OR "RCC"?

```{r}
sample_meta[sample_meta$indication %in% c("CRC", "HNSCC", "BLCA", "RCC"),]
```

in sample_meta, can you extract out the row with rowname
"X7465.DA.0086"?

```{r}
sample_meta["X7465.DA.0086",]
```

in sample_meta, can you subset out the column with colname "indication"?

```{r}
sample_meta[,"indication"]
```

in the gene_features table, can you extract out the row with ensembl_id
= "ENSG00000002586"?

```{r}
gene_features[gene_features$ensembl_gene_id == "ENSG00000002586",]
```

in rna_counts, can you extract out the rows (genes) which have values \>
1000 for the sample X7465.DA.0012?

```{r}
rna_counts[rna_counts$X7465.DA.0012 > 1000,]
```

what is the data type of 1) sample_meta's "indication" column 2)
gene_features' "" column and 3) rna counts' "X7465.DA.0086" column?

```{r}
class(sample_meta$indication)
class(rna_counts$X7465.DA.0086)
```

Can you convert sample_meta's "indication" column into the data type
factor?

```{r}
as.factor(sample_meta$indication)
```

Can you convert it back to what it was originally?

```{r}
as.vector(factor(sample_meta$indication))
```

In sample_meta, can you tally up how many samples we have for each
indication? (hint: use the built-in "table" function)

```{r}
table(sample_meta$indication)
```

#### 2) Make DGE list object and include sample/ gene feature info

```{r}
y <- DGEList(rna_counts, samples = sample_meta, genes = gene_features)
```

Let's play around with this object, y, a little: Questions: Can you see
what slots are stored in y?

```{r}
names(y)
```

Can you print out the first 6 rows of "samples" information in y?

```{r}
y$samples[1:6,]
```

Can you write some code to print out just the library sizes in y?

```{r}
print(y$samples$lib.size)
```

Can you calculate 1) the sum of library sizes of all samples 2) the mean
library size?

```{r}
sum(y$samples$lib.size)
mean(y$samples$lib.size)
```

Can you make a barplot of library sizes for all the samples?

```{r}
barplot(y$samples$lib.size, main = "Sample Library Sizes", ylab = "Library Size")
```

Do you notice any potentially problematic samples?

```{r}
y$samples[1,] # library size is low for sample 1
```
