### Setup

```{r}
library(edgeR)
library(limma)
library(RColorBrewer)
library(gplots)
library(ggbiplot)
library(pheatmap)
library(reshape2)
library(matrixStats)
library(tidyverse)
```

### Data

Using log2 expression data with only protein-coding genes. 'Reduced Location' refers to the location of the cancer but with certain categories combined: 'Larynx', 'Nasopharynx', 'Oropharynx', and 'Oral Cavity' categorites are now 'Head/Neck'; 'Colon' and 'Colon/Rectum' categories are now 'Colon/Rectum'; and 'Lung' and 'Lung & Bronchus' categories are now 'Lung/Bronchus'.

```{r}
expset <- readRDS('../data/virus_expset.rds')
expressed <- expset %>%
                       .[rowSums(exprs(.)) > 0] %>%
                       fData() %>%
                       .$RefSeq %>%
                       unique()
expset <- expset[fData(expset)$RefSeq %in% expressed]
expset <- expset[order(rownames(fData(expset)))]
expset <- expset[,order(pData(expset)$reduced_location)]
locations <- pData(expset)$reduced_location
```

### Heatmap

```{r}
ann <- data.frame(location=locations)
rownames(ann) <- colnames(expset)
ann_colors <- list(location=locations %>%
                              unique() %>%
                              setNames(brewer.pal(length(.), 'Paired'), .))

pdf('../output/virus_expressed_genes_heatmaps.pdf')

for (gene in expressed) {
  gene_symbol <- fData(expset) %>% 
                   .[.$RefSeq == gene,] %>% 
                   .$Abbreviation
  gene_num <- gene_symbol %>% length()
  gene_symbol <- gene_symbol %>% .[1]
  for (scale in c('row', 'none')) {
    pheatmap(exprs(expset)[fData(expset)$RefSeq == gene,], 
             scale=scale,
             cluster_rows=F, 
             cluster_cols=F,
             # show_colnames=F,
             fontsize=2,
             fontsize_row=4,
             fontsize_column=2,
             annotation_col=ann,
             annotation_colors=ann_colors,
             cellheight=min(460 / gene_num, 20),
             gaps_col=match(unique(locations), locations) - 1,
             labels_row=fData(expset)$hgnc_symbol,
             main=str_glue('{gene_symbol} logTPM by sample (Scaled by {scale})'))
  }
}

dev.off()
```

```{r}
specific <- c('lcl|NC_001355.1_cds_NP_040296.1_1',
              'lcl|NC_001355.1_cds_NP_040297.1_2',
              'lcl|NC_001355.1_cds_NP_040298.1_3',
              'lcl|NC_001355.1_cds_NP_040299.1_4',
              'lcl|NC_001355.1_cds_NP_040300.1_5',
              'lcl|NC_001355.1_cds_NP_040301.1_6',
              'lcl|NC_001355.1_cds_NP_040302.1_7',
              'lcl|NC_001355.1_cds_NP_040303.1_8',
              'lcl|NC_001355.1_cds_NP_040304.1_9',
              'lcl|NC_006273.2_cds_YP_081567.1_115',
              'lcl|NC_006273.2_cds_YP_081597.1_146',
              'lcl|NC_007605.1_cds_YP_401645.1_15',
              'lcl|NC_007605.1_cds_YP_401654.1_25',
              'lcl|NC_007605.1_cds_YP_401684.3_53',
              'lcl|NC_007605.1_cds_YP_401695.1_66',
              'lcl|NC_007605.1_cds_YP_401707.1_79',
              'lcl|NC_007605.1_cds_YP_401712.1_84',
              'lcl|NC_007605.1_cds_YP_401713.1_85')

data <- expset %>%
          exprs() %>%
          .[specific,]
          # .[(fData(expset)$RefSeq %in% specific) & rowSums(.) > 0,]

View(expressed)

pheatmap(exprs(expset)[fData(expset)$RefSeq == 'NC_001355',pData(expset)$reduced_location=='Head/Neck'], 
             scale=scale,
             cluster_rows=F, 
             cluster_cols=F,
             fontsize=9,
             fontsize_row=4,
             annotation_col=ann,
             annotation_colors=ann_colors,
             cellheight=min(460 / gene_num, 20),
             labels_row=fData(expset)$hgnc_symbol,
             main=str_glue('{gene_symbol} logTPM by sample (Scaled by {scale})'))

```
